Comment: Reasoning about Interference Between Units

### 2015-12-23 response.md

In the paper we clarified that we are focusing on testing sharp null
hypotheses arising from one toy model of how an experimental treatment might
propagate across a social network. We have added some discussion to the paper
highlighting the fact that the power of a test of a sharp null hypothesis
depends on *both* the model of causal effects *and* the test statistic (and
the effective sample size which itself depends on sample size and network
characteristics). We also note that, given randomization, any test statistic
ought to produce valid statistical inferences (i.e. tests that would reject
the truth no more than $100 \alpha \%$ of the time).

Since the point of the paper is to learn about test statistics,
we do not explore alternative models of causal effects. The original
paper (BFP) on this method compared the performance of tests when the data did
and did not arise from the causal model and we point to that paper for those
who would like to pursue those set of questions. We did add some discussion
about the power of the SSR test statistic against different kinds of causal
effects: the SSR test statistic ought to be particularly sensitive to changes
in mean differences (after all, recall that the coefficient on a dichotomous
variable in an least squares model with a continuous outcome that minimizes
the SSR is the difference of means) and ought to be less sensitive to other
differences.

2) Reviewer 2 asks you to carry forward your toy example with four units
through the entire analysis.  If this isn't possible, please provide the
simplest case that illuminate the core intuition of your method.

When we set about adding the four unit example to the different sections of the
paper, we realized that we were not enhancing intuition: the two test
statistics applied to an experiment of size 4 mainly reflected the fact that we
had very little data. Instead, we added a discussion of the SSR and KS test
statistics applied to a case with no interference and with very simple models
of effect. We also added a figure and discussion that tries to show a bit more
about how models of effects work in the context of sharp null hypothesis
testing. We hope that this helps with the intuition so that the use of the SSR
test statistic to an experiment on a social network with a complex model of
treatment propagation is easier to understand.

### 2015-12-23 paper/introduction.tex
Working on the 'intuition' of SSR versus KS part
Added section describing the simple SSR versus KS comparison

\begin{figure}[H]\centering
  \includegraphics[width=.5\textwidth]{figures/ksvsssroutcomes.pdf}
  \caption{}\label{fig:simpleoutcomes}
\end{figure}

And we compared two models of effects --- a constant additive
effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant
multiplicative effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set
the truth to be the sharp null of no effects $\tau=0$ for the additive model
and $\tau=1$ for the multiplicative model. Our intuition was that the SSR
would have more power than the KS test when the outcome was Normal and when
the observational implication of the model of effects would shift means (i.e.
the additive model). We used direct simulation of the randomization
distribution to generate $p$-values and repeated that process 1000 times to
gauge the proportion of rejections of false hypotheses (i.e. the power of the
tests). The results, not shown here but part of the reproduction archive for
this paper, bear out this intuition: the SSR has slightly more power than KS
for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

In the case where we know
the fixed adjacency matrix of the network, $\bS$, and where we imagine that
network attributes (like degree) of a node play a role in the mechanism by
which treatment propagates, the idea of assessing model fit rather than
closeness of distributions leads naturally to the sum-of-squared-residuals
(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
coefficients $\bbeta$ when we could define the test statistic as:

>>>>

\begin{figure}[H]\centering
  \includegraphics[width=.5\textwidth]{ksvsssr-outcomes.pdf}
  \caption{Two outcomes ($Y_i$) simulated with no interference, $N=256$, half
    randomized to binary treatment by complete randomization. }\label{fig:simpleoutcomes}
\end{figure}

And we compared two models of effects --- a constant additive effects model in
which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant multiplicative effects
model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set the truth to be the
sharp null of no effects $\tau=0$ for the additive model and $\tau=1$ for the
multiplicative model. To further explain the process of hypothesis testing and
evaluation of the test statistics, we display the results from a power
analysis of one alternative hypothesis in figure~\ref{fig:boxplot}. A given
model and hypothesized parameter implies a distribution for observed outcomes
in the treatment group and control groups. Here, since we have no
interference, the hypotheses only have different implications for the control
group distribution. In this case, we chose a two parameters for the additive
and multiplicative models which produced very similar implications (as show by
the similarity in the ``Add. Model'' and ``Mult.  Model'' boxplots in both
panels of the figure). We see that when we tested the hypothesis of
$\tau_0=.34$ from the additive model using the KS test statistic on Normal
data, that we produced $p$-values lower than $\alpha=.05$ about 55\% of the
time across 1000 simulations. That is, the KS test has a power of about .55
to reject a false hypothesis with this Normal outcome and this constant
additive effects model. The analogous power for the SSR test statistic was
.73. For these particular parameters, we see the SSR performing better with
Normal outcomes but not being very effected by the model of effects (which
makes some sense here because we are choosing parameters at which both models
imply more or less the same patterns in outcomes).

\begin{figure}[H]\centering
  \includegraphics[width=.5\textwidth]{ksvsssr-boxplot.pdf}
  \caption{Distributions of the simulated data from left to right: $y_0$
    outcome with no experiment, observed outcome after random assignment,
    outcomes implied by a constant additive model of effects, outcomes implied by a constant
    multiplicative model. The hypothesized model parameters $\tau_0$ that
    produce the patterns shown are printed on the plot. The proportion of
    simulated tests in which these values of $\tau_0$ with $p$-values less
    than $\alpha=.05$ by the KS and SSR test statistics are printed on the plot as $\text{pow}_{KS}$ and
  $\text{pow}_\text{SSR}$. The SSR test statistic has more power for the
  Normal outcomes and less power for the skewed outcome.}\label{fig:boxplot}
\end{figure}

When we assessed power across many alternative hypotheses, our intuition was that the SSR
would have more power than the KS test when the outcome was Normal and when
the observational implication of the model of effects would shift means (i.e.\
the additive model). We used direct simulation of the randomization
distribution to generate $p$-values and repeated that process 1000 times to
gauge the proportion of rejections of false hypotheses (i.e.\ the power of the
tests). The results, not shown here but part of the reproduction archive for
this paper, bear out this intuition: the SSR has slightly more power than KS
for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

\subsection{The SSR Test Statistic with Network Information}

In the case where we know the fixed adjacency matrix of a social network,
$\bS$, and where we imagine that network attributes (like degree) of a node
play a role in the mechanism by which treatment propagates, the idea of
assessing model fit rather than closeness of distributions leads naturally to
the sum-of-squared-residuals (SSR) from a least squares regression of $
\widetilde y_{i,Z_i=0}$ on $Z_{i}$ and $\bz^{T} \bS$ (i.e.\ the number of
directly connected nodes assigned treatment) as well as the $\mathbf{1}^{T}
\bS$ (i.e.\ the degree of the node). If we collected $Z_{i}$,$\bz^{T} \bS$,
and $\mathbf{1}^{T} \bS$ into a matrix $\bX$, and fit the $\widetilde
y_{i,Z_i=0}$ as a linear function of $\bX$ with coefficients $\bbeta$ when we
could define the test statistic as:
